
```{r}
# install.packages("tidyverse")
library(tidyverse)

# Assuming your data frame is called HeartFailure
HeartFailure <- HeartFailure %>%
  mutate(DEATH_EVENT = ifelse(DEATH_EVENT == 1, "Dead", "Alive"))

HeartFailure$DEATH_EVENT <- factor(HeartFailure$DEATH_EVENT)
```

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Cmd+Shift+Enter*. 

```{r}
# importing and load package
install.packages("openxlsx")
library(openxlsx)

#load dataset
#data <- read.xlsx(HeartFailure.xlsx)
```

1. Visualization

(a.) Develop an overall view of relationship of dependent variable (Death event) with
all continuous variables (see iris example with scatterplot matrix in Rcmdr)
```{r}

```

b. Develop at least two perceptrons where there are three or more x-variables and
Death event is the y-variable; calculate a measure of accuracy for each
```{r}

```

c. Show at least two graphs with pairs of continuous variables and plot Death
events (see iris example in RStudio) (10 points)
```{r}
ggplot(HeartFailure, aes(x = age, y = platelets)) +
geom_point(aes(colour = DEATH_EVENT, shape= DEATH_EVENT), size = 3) +
xlab("Age") +  # Corrected x-axis label
ylab("platelets") +  # Corrected y-axis label
ggtitle("Death event vs Age and platelets") +
theme(plot.title = element_text(hjust = 0.5))
```

```{r}
ggplot(HeartFailure, aes(x = serum_sodium, y = serum_creatinine)) +
geom_point(aes(colour = DEATH_EVENT, shape= DEATH_EVENT), size = 3) +
xlab("Serum Sodium") + # Corrected x-axis label
ylab("Serum Creatinine") + # Corrected y-axis label
ggtitle("Serum Sodium vs Serum Creatinine") +
theme(plot.title = element_text(hjust = 0.5))
```
```{r}
set.seed(7) # We went for this because it is your lucky number

HeartFailure$DEATH_EVENT <- lapply(HeartFailure$DEATH_EVENT, function(x) {
if(x == 1)
HeartFailure$DEATH_EVENT <- -1
else if(x == 0)
HeartFailure$DEATH_EVENT <- 1
else
HeartFailure$DEATH_EVENT <- NULL
})
```

2. Perceptrons (25 points)

a. Develop at least two perceptrons where there are two x-variables, using continuous and binary x-variables, with Death event as the y-variable; calculate a measure of accuracy for each (10 points)

```{r}
#Creating training and testing data
HeartFailure_index <- sample(nrow(HeartFailure), 0.7*nrow(HeartFailure))
HeartFailure_train <- HeartFailure[HeartFailure_index, ]
HeartFailure_test <- HeartFailure[-HeartFailure_index, ]
```


```{r}
# Create the Perceptron Learning Algorithm
perceptron <- function(X, y, numEpochs) {
  results <- list()
  w <- runif(ncol(X), -10, 10) #Initalize weights

# For loop - number of generations(epochs) - number of times dataset is ran through
for(j in 1:numEpochs) {
  predictedResult <- numeric(length=100) # Initalize predictedResult vector
  numIncorrect = 0 # Keeps track of # of missclassified points

  # For loop - loop throught dataset
  for(i in 1:length(y)) {
    xi = as.numeric(unlist(X[i,])) # Convert dataframe to vector
    predictedResult[i] = sign(w %*% xi) # Predict the point
  
    # If predicted point is incorrect - change weight
    if(predictedResult[i] != y[i]) {
      numIncorrect = numIncorrect + 1 # Add one to # of missclassified points
      w <- w + as.numeric(y[i]) * xi # Update the weight w <- w + WiXi
    }
  }
  # Print results of this generation(epoch)
  cat("\nEpoch #: ", j)
  cat("\nNumber Incorrect: ", numIncorrect)
  cat("\nFinal Weight: ", w)
 }
}  
```

```{r}
set.seed(7)
#Creating first Perceptron model
X1 <- HeartFailure_train[, c("smoking","high_blood_pressure")]
Y1 <- HeartFailure_train$DEATH_EVENT
perceptron(X1,Y1,8)

#Take the weight from last Epoch
weight = c(97.77819, 94.95491)

#Multiply weights to Smoking and High Blood Pressure
HeartFailure_test$Wsmoking <- HeartFailure_test$smoking*weight[1]
HeartFailure_test$WHBP <- HeartFailure_test$high_blood_pressure*weight[2]

#Add all weights and call it predict
HeartFailure_test$predict <- rowSums(HeartFailure_test[,c("Wsmoking","WHBP")])

#Use confusion matrix to evaluate the model performance
perceptron1predict <- table(HeartFailure_test$DEATH_EVENT == 1, HeartFailure_test$predict > 0) + 
  table(HeartFailure_test$DEATH_EVENT == 0, HeartFailure_test$predict == 0)

perceptron1predict


#Get the accuracy rate
sum(diag(perceptron1predict))/sum(perceptron1predict)
```

```{r}
\

#Take the weight from last Epoch
weight = c(9.778186, 0.9549091)

#Multiply weights to Age and Platelets
HeartFailure_test$Wanaemia <- HeartFailure_test$anaemia*weight[1]
HeartFailure_test$Wdiabetes <- HeartFailure_test$diabetes*weight[2]

#Add all weights and call it predict
HeartFailure_test$predict <- rowSums(HeartFailure_test[,c("Wanaemia","Wdiabetes")])

#Use confusion matrix to evaluate the model performance
perceptron4predict <- table(HeartFailure_test$DEATH_EVENT == 1, HeartFailure_test$predict > 0) + 
  table(HeartFailure_test$DEATH_EVENT == 0, HeartFailure_test$predict == 0)

perceptron4predict


#Get the accuracy rate
sum(diag(perceptron4predict))/sum(perceptron4predict)
```
```{r}
set.seed(7)
#Creating first Perceptron model
X5 <- HeartFailure_train[, c("age","platelets")]
Y5 <- HeartFailure_train$DEATH_EVENT
perceptron(X5,Y5,8)

#Take the weight from last Epoch
weight = c(81.77819, 620998)

#Multiply weights to Age and Platelets
HeartFailure_test$Wage <- HeartFailure_test$age*weight[1]
HeartFailure_test$Wplatelets <- HeartFailure_test$platelets*weight[2]

#Add all weights and call it predict
HeartFailure_test$predict <- rowSums(HeartFailure_test[,c("Wage","Wplatelets")])

#Use confusion matrix to evaluate the model performance
perceptron5predict <- table(HeartFailure_test$DEATH_EVENT == 1, HeartFailure_test$predict > 0) + 
  table(HeartFailure_test$DEATH_EVENT == 0, HeartFailure_test$predict == 0)

perceptron5predict


#Get the accuracy rate
sum(diag(perceptron5predict))/sum(perceptron5predict)
```




```{r}
# Create training and testing data
# Separate the data into training and testing. Selecting 70% of data for training dataset, 30% for testing dataset.

HeartFailure_index <- sample(nrow(HeartFailure), 0.7 * nrow(HeartFailure)) 
HeartFailure_train <- HeartFailure[HeartFailure_index, ]
HeartFailure_test <- HeartFailure[-HeartFailure_index, ]
```

```{r}
# First perceptron model
# Using train data set for building perceptron model
#Separate the data into input (X variables) and output (Y variable)

X <- HeartFailure_train[, c("age", "smoking")] # Input Matrix 
y <- HeartFailure_train$DEATH_EVENT # Output Vector
```

```{r}
# Run the Perceptron algorithm
perceptron(X,y, 8)
```

```{r}
# Interpretation

#Take the weight from last Epoch (lowest Number Incorrect)
weight = c(27.0542, 9.883745)

#Have age and smoking multipling weight accordingly
HeartFailure_test$Wage <- HeartFailure_test$age*weight[1] 
HeartFailure_test$Wsmoking <- HeartFailure_test$smoking*weight[2]

# Add all calculated weights, and name it as predict
HeartFailure_test$predict <- rowSums(HeartFailure_test[,c("Wage","Wsmoking")])
```

```{r}
# Use confusion matrix to evaluate the model performance
perceptronpredicttable <- table(HeartFailure_test$DEATH_EVENT == 1, HeartFailure_test$predict > 0) + table(HeartFailure_test$DEATH_EVENT == -1, HeartFailure_test$predict <0)

perceptronpredicttable
```
```{r}
# get the accuracy rate
sum(diag(perceptronpredicttable))/sum(perceptronpredicttable)
```

```{r}
# Second perceptron model
# Create training and testing data
# Separate the data into training and testing. Selecting 70% of data for training dataset, 30% for testing dataset.

set.seed(7)

HeartFailure_index2 <- sample(nrow(HeartFailure), 0.7 * nrow(HeartFailure)) 
HeartFailure_train2 <- HeartFailure[HeartFailure_index2, ]
HeartFailure_test2 <- HeartFailure[-HeartFailure_index2, ]
```

```{r}
# Second perceptron model
# Using train data set for building perceptron model
#Separate the data into input (X variables) and output (Y variable)

XX <- HeartFailure_train2[, c("platelets", "high_blood_pressure")] # Input Matrix 
yy <- HeartFailure_train2$DEATH_EVENT # Output Vector
```

```{r}
# Run the Perceptron algorithm
perceptron(XX,yy, 8)
```

```{r}
# Interpretation

#Take the weight from last Epoch (lowest Number Incorrect)
weight = c(0.0542027, 8.883745)

#Have age and smoking multipling weight accordingly
HeartFailure_test2$Wplatelets <- HeartFailure_test2$platelets*weight[1] 
HeartFailure_test2$Whigh_blood_pressure <- HeartFailure_test2$high_blood_pressure*weight[2]

#HeartFailure_test$Wsmoking <- HeartFailure_test$smoking*weight[1]
#HeartFailure_test$WHBP <- HeartFailure_test$high_blood_pressure*weight[2]

# Add all calculated weights, and name it as predict
HeartFailure_test2$predict <- rowSums(HeartFailure_test2[,c("Wplatelets","Whigh_blood_pressure")])
```

```{r}
# Use confusion matrix to evaluate the model performance
perceptron1predicttable <- table(HeartFailure_test2$DEATH_EVENT == 1, HeartFailure_test2$predict > 0) + table(HeartFailure_test2$DEATH_EVENT == 0, HeartFailure_test2$predict == 0)

perceptron1predicttable
```

```{r}
# get the accuracy rate
sum(diag(perceptron1predicttable))/sum(perceptron1predicttable)

```

b. Develop at least two perceptrons where there are three or more x-variables and Death event is the y-variable; calculate a measure of accuracy for each
(10 points)

```{r}
# First perceptron model

# Create training and testing data
# Separate the data into training and testing. Selecting 70% of data for training dataset, 30% for testing dataset.
HeartFailure_index3 <- sample(nrow(HeartFailure), 0.7 * nrow(HeartFailure)) 
HeartFailure_train3 <- HeartFailure[HeartFailure_index3, ]
HeartFailure_test3 <- HeartFailure[-HeartFailure_index3, ]

# Using train data set for building perceptron model
#Separate the data into input (X variables) and output (Y variable)
XXX <- HeartFailure_train3[, c("age","creatinine_phosphokinase","serum_creatinine", "serum_sodium", "platelets", "time")] # Input Matrix 
yyy <- HeartFailure_train3$DEATH_EVENT # Output Vector

# Run the Perceptron algorithm
perceptron(XXX, yyy, 8)

# Interpretation
#Take the weight from last Epoch (lowest Number Incorrect)
weight = c(2.204697, -9.281745, 4.551681, -9.339275, 5.900778, 8.351514)

#Have age and smoking multipling weight accordingly
HeartFailure_test3$Wplatelets <- HeartFailure_test3$age*weight[1] 
HeartFailure_test3$Wcreatinine_phosphokinase <- HeartFailure_test3$creatinine_phosphokinase*weight[2]
HeartFailure_test3$Wserum_creatinine <- HeartFailure_test3$serum_creatinine*weight[3]
HeartFailure_test3$Wserum_sodium <- HeartFailure_test3$serum_sodium*weight[4]
HeartFailure_test3$Wplatelets <- HeartFailure_test3$platelets*weight[5]
HeartFailure_test3$Wtime <- HeartFailure_test3$time*weight[6]

# Add all calculated weights, and name it as predict
HeartFailure_test3$predict <- rowSums(HeartFailure_test3[,c("Wplatelets","Wcreatinine_phosphokinase", "Wserum_creatinine", "Wserum_sodium","Wplatelets", "Wtime")])

# Use confusion matrix to evaluate the model performance
perceptronpredicttable3 <- table(HeartFailure_test3$DEATH_EVENT == 1, HeartFailure_test3$predict > 0) + table(HeartFailure_test3$DEATH_EVENT == 0, HeartFailure_test3$predict == 0)

perceptronpredicttable3

# get the accuracy rate
sum(diag(perceptronpredicttable3))/sum(perceptronpredicttable3)
```

```{r}
# First perceptron model ka alternative
# Create training and testing data
# Separate the data into training and testing. Selecting 70% of data for training dataset, 30% for testing dataset.

HeartFailure_index3a <- sample(nrow(HeartFailure), 0.7 * nrow(HeartFailure)) 
HeartFailure_train3a <- HeartFailure[HeartFailure_index3a, ]
HeartFailure_test3a <- HeartFailure[-HeartFailure_index3a, ]
```

```{r}
# Using train data set for building perceptron model
#Separate the data into input (X variables) and output (Y variable)
age + creatinine_phosphokinase + serum_creatinine + serum_sodium + platelets + time
XXXa <- HeartFailure_train3[, c("age","high_blood_pressure","sex","smoking")] # Input Matrix 
yyya <- HeartFailure_train3$DEATH_EVENT # Output Vector
```


```{r}
# Run the Perceptron algorithm
perceptron(XXXa, yyya, 8)
```

```{r}
# Interpretation

#Take the weight from last Epoch (lowest Number Incorrect)
weight = c(3.347024, -2.747504, 2.044756, -7.920739)

#Have age and smoking multipling weight accordingly
HeartFailure_test3$Wplatelets <- HeartFailure_test3$age*weight[1] 
HeartFailure_test3$Whigh_blood_pressure <- HeartFailure_test3$high_blood_pressure*weight[2]
HeartFailure_test3$Wsex <- HeartFailure_test3$sex*weight[2]
HeartFailure_test3$Wsmoking <- HeartFailure_test3$smoking*weight[2]

# Add all calculated weights, and name it as predict
HeartFailure_test3$predict <- rowSums(HeartFailure_test3[,c("Wplatelets","Whigh_blood_pressure", "Wsex", "Wsmoking")])
```

```{r}
# Use confusion matrix to evaluate the model performance
perceptronpredicttable3a <- table(HeartFailure_test3$DEATH_EVENT == 1, HeartFailure_test3$predict > 0) + table(HeartFailure_test3$DEATH_EVENT == 0, HeartFailure_test3$predict == 0)

perceptronpredicttable3a
```

```{r}
# get the accuracy rate
sum(diag(perceptronpredicttable3a))/sum(perceptronpredicttable3a)
```

```{r}
# Second perceptron model
# Create training and testing data
# Separate the data into training and testing. Selecting 70% of data for training dataset, 30% for testing dataset.

HeartFailure_index4 <- sample(nrow(HeartFailure), 0.7 * nrow(HeartFailure)) 
HeartFailure_train4 <- HeartFailure[HeartFailure_index4, ]
HeartFailure_test4 <- HeartFailure[-HeartFailure_index4, ]
```

```{r}
# Second perceptron model
# Using train data set for building perceptron model
#Separate the data into input (X variables) and output (Y variable)

X <- HeartFailure_train4[, c("age", "anaemia","creatinine_phosphokinase", "diabetes", "ejection_fraction","platelets", "high_blood_pressure")] # Input Matrix 
y <- HeartFailure_train4$DEATH_EVENT # Output Vector
```

```{r}
# Create the Perceptron Learning Algorithm
perceptron <- function(X, y, numEpochs) {
  results <- list()
  w <- runif(ncol(X), -10, 10) #Initalize weights

# For loop - number of generations(epochs) - number of times dataset is ran through
for(j in 1:numEpochs) {
  predictedResult <- numeric(length=100) # Initalize predictedResult vector
  numIncorrect = 0 # Keeps track of # of missclassified points

  # For loop - loop throught dataset
  for(i in 1:length(y)) {
    xi = as.numeric(unlist(X[i,])) # Convert dataframe to vector
    predictedResult[i] = sign(w %*% xi) # Predict the point
  
    # If predicted point is incorrect - change weight
    if(predictedResult[i] != y[i]) {
      numIncorrect = numIncorrect + 1 # Add one to # of missclassified points
      w <- w + as.numeric(y[i]) * xi # Update the weight w <- w + WiXi
    }
  }
  # Print results of this generation(epoch)
  cat("\nEpoch #: ", j)
  cat("\nNumber Incorrect: ", numIncorrect)
  cat("\nFinal Weight: ", w)
 }
}  
```

```{r}
# Run the Perceptron algorithm
perceptron(X,y, 8)
```

```{r}
# Interpretation

#Take the weight from last Epoch (lowest Number Incorrect)
weight = c(-2.802817, 2.349683, 4.094498, -3.908247, -6.899839, 9.054003, 7.058395)

#Have age and smoking multipling weight accordingly
HeartFailure_test4 $Wage <- HeartFailure_test4$age*weight[1] 
HeartFailure_test4$Wanaemia <- HeartFailure_test4$anaemia*weight[2]
HeartFailure_test4$Wcreatinine_phosphokinase <- HeartFailure_test4$creatinine_phosphokinase*weight[3]
HeartFailure_test4$Wdiabetes <- HeartFailure_test4$diabetes*weight[4]
HeartFailure_test4$Wejection_fraction <- HeartFailure_test4$ejection_fraction*weight[5]
HeartFailure_test4$Wplatelets <- HeartFailure_test4$platelets*weight[6]
HeartFailure_test4$Whigh_blood_pressure <- HeartFailure_test4$high_blood_pressure*weight[7]

# Add all calculated weights, and name it as predict
HeartFailure_test4$predict <- rowSums(HeartFailure_test4[,c("Wplatelets","Whigh_blood_pressure")])
```

```{r}
# Use confusion matrix to evaluate the model performance
perceptronpredicttable4 <- table(HeartFailure_test4$DEATH_EVENT == 1, HeartFailure_test4$predict > 0) + table(HeartFailure_test4$DEATH_EVENT == 0, HeartFailure_test4$predict == 0)

perceptronpredicttable4
```

```{r}
# get the accuracy rate
sum(diag(perceptronpredicttable4))/sum(perceptronpredicttable4)
```


c. Summarize the accuracy of all the perceptron models (5 points)
```{r}

```

3. Support Vector Machines (SVM) (35 points)
a. Develop at least two SVMs where there are two x-variables, using continuous and binary x-variables, with Death event as the y-variable; calculate a measure of accuracy for each (10 points)
```{r}
age+high_blood_pressure+sex+smoking
```

```{r}
# load svm package
library(e1071)
HeartFailure$DEATH_EVENT <- as.factor(HeartFailure$DEATH_EVENT)
```

```{r}
# First SVM
#To run the SVM,
model1 <- svm(DEATH_EVENT ~ age + platelets, data = HeartFailure)

#To print the model and a summary, type
print(model1)
summary(model1)

#To display the model, type
#plot(model1,HeartFailure)
plot(model1, data = HeartFailure, age ~ platelets)

#Predict the model
heartpredict1 <- predict(model1,HeartFailure)
predicttable1 <- table(heartpredict1, HeartFailure$DEATH_EVENT)
predicttable1

#Calculate the accuracy
sum(diag(predicttable1))/sum(predicttable1)

# Prediction
#predicteda <- predict(model1, HeartFailure)

# Create a confusion matrix
#confusionMatrix(predicteda, HeartFailure$DEATH_EVENT)
```

```{r}
# Second SVM
#To run the SVM,
model2 <- svm(DEATH_EVENT ~ age + time, data = HeartFailure)

#To print the model and a summary, type
print(model2)
summary(model2)

#To display the model, type
plot(model2, data = HeartFailure, age ~ time)

#Predict the model
heartpredict2 <- predict(model2,HeartFailure)
predicttable2 <- table(heartpredict2, HeartFailure$DEATH_EVENT)
predicttable2

#Calculate the accuracy
sum(diag(predicttable2))/sum(predicttable2)

# Prediction
#predictedb <- predict(model2, HeartFailure)

# Create a confusion matrix
#confusionMatrix(predicteda, HeartFailure$DEATH_EVENT)
```

```{r}
# Second SVM' alternative
#To run the SVM,
model3 <- svm(DEATH_EVENT ~ ejection_fraction + age, data = HeartFailure)

#To print the model and a summary, type
print(model3)
summary(model3)

#To display the model, type
plot(model3, data = HeartFailure, ejection_fraction ~ age)

# Prediction
#predictedc <- predict(model3, HeartFailure)

# Create a confusion matrix
#confusionMatrix(predicteda, HeartFailure$DEATH_EVENT)

```

```{r}
# Second SVM' alternatives' alternative
#To run the SVM,
model4 <- svm(DEATH_EVENT ~ ejection_fraction + platelets, data = HeartFailure)

#To print the model and a summary, type
print(model4)
summary(model4)

#To display the model, type
plot(model4, data = HeartFailure, ejection_fraction ~ platelets)

# Prediction
predicteda <- predict(model4, HeartFailure)

# Create a confusion matrix
confusionMatrix(predicteda, HeartFailure$DEATH_EVENT)
```

b. Develop at least two SVMs where there are three or more x-variables, with Death event as the y-variable; calculate a measure of accuracy for each (10 points)
```{r}
# First SVM
#To run the SVM,
model6 <- svm(DEATH_EVENT ~ age + creatinine_phosphokinase + ejection_fraction, data = HeartFailure)

#To print the model and a summary, type
print(model6)
summary(model6)

#To display the model, type
#plot(model6, data = HeartFailure, age ~ creatinine_phosphokinase ~ ejection_fraction)

predicted <- predict(model6, HeartFailure)

# Create a confusion matrix
confusionMatrix(predicted, HeartFailure$DEATH_EVENT)
```

```{r}
# Second SVM
#To run the SVM,
model7 <- svm(DEATH_EVENT ~ age + high_blood_pressure + sex + smoking, data = HeartFailure)

#To print the model and a summary, type
print(model7)
summary(model7)

#To display the model, type
#plot(model6, data = HeartFailure, age ~ high_blood_pressure ~ sex ~ smoking)

predicted2 <- predict(model7, HeartFailure)

# Create a confusion matrix
confusionMatrix(predicted2, HeartFailure$DEATH_EVENT)
```

```{r}
# Second SVM ka alternative
#To run the SVM,
model7a <- svm(DEATH_EVENT ~ age + high_blood_pressure + sex + smoking + time, data = HeartFailure)

#To print the model and a summary, type
print(model7a)
summary(model7a)

#To display the model, type
#plot(model6, data = HeartFailure, age ~ high_blood_pressure ~ sex ~ smoking)

predicted2 <- predict(model7a, HeartFailure)

# Create a confusion matrix
confusionMatrix(predicted2, HeartFailure$DEATH_EVENT)
```

```{r}
# Third SVM
# To run the SVM,
model8 <- svm(DEATH_EVENT ~ age + creatinine_phosphokinase + serum_creatinine + serum_sodium + platelets + time, data = HeartFailure)

#To print the model and a summary, type
print(model8)
summary(model8)

#To display the model, type
#plot(model8, data = HeartFailure, age ~ creatinine_phosphokinase ~ serum_creatinine ~ serum_sodium ~ platelets)

predicted3 <- predict(model8, HeartFailure)

# Create a confusion matrix
confusionMatrix(predicted3, HeartFailure$DEATH_EVENT)
```

c. Generate a graphic showing the results of each of the SVM results (10 points)
```{r}

```

d. Summarize the accuracy of all the SVM models (5 points)
```{r}

```


Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

